---
title: "Inferential Network Statistics"
description: |
  A closer look into the S&P 500 stock network properties.
author:
  - name: Peter Sullivan
    url: {}
date: 2022-04-24
preview: preview.png
output:
  distill::distill_article:
    self_contained: false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning =  FALSE)
library(tidyverse)
library(tidyquant)
library(igraph)
library(knitr)
library(statnet)


small_network <- readRDS("Smallnetwork.RDS")


```


This is the same network that I used for Hw 7. The small network above is a filtered correlation network created from the S&P 500. I grabbed the close prices from all s&p 500 stocks for 2021. I then created a correlation matrix from those prices. I then created ties for stocks that were 98% correlated. This filtered down the list of 500 stocks to 32 stocks. I then filtered out any isolates.


## CUG-tests to Test Network Properties


I first need to change my network objects into adjacency matrixs.

```{r}


small.mat <- as.matrix(as_adjacency_matrix(small_network))



small.cug <- sna::cug.test(small.mat,FUN=gtrans,mode="digraph",cmode="size", reps = 100)


small.cug
plot(small.cug)


paste("The T stat: ",(small.cug$obs.stat-mean(small.cug$rep.stat))/sd(small.cug$rep.stat))

```

There is a low probability the transitivity can be randomly generated.

## Creating T stat funciton

```{r}
cug.t<-function(cug.object){
  (cug.object$obs.stat-mean(cug.object$rep.stat))/sd(cug.object$rep.stat)
}
```



## Cug test to different Network Properties

### Degree

```{r}
small.cug <-cug.test(small.mat,FUN=centralization,  FUN.arg=list(FUN=degree, cmode="indegree"), mode="digraph", cmode="size") 
#plot vs simulation results
plot(small.cug)

paste("T Statistic: ",cug.t(small.cug))

```
There is a good probability that the degree could be randomly generated.


### Betweenness

```{r}


small.bet <-cug.test(small.mat,FUN=centralization,  FUN.arg=list(FUN=betweenness, cmode="directed"), mode="digraph", cmode="size", reps=100) 
#plot vs simulation results
plot(small.bet)

paste("T stat: ", cug.t(small.bet))
```
There is a very small probability that the betweeness could be randomly generated.


### Transitivity

```{r}

small.trans <- cug.test(small.mat,FUN=gtrans,mode="digraph",cmode="dyad")
small.trans
plot(small.trans)
paste("T stat: ", cug.t(small.trans))
```

### Edges

```{r}

small.edge<-cug.test(small.mat,FUN=gtrans,mode="digraph",cmode="edges", reps=100)
small.edge
plot(small.edge)
paste("T stat: ", cug.t(small.edge))

```

